All Graphs Lead to Rome: Learning Geometric and Cycle-Consistent Representations with Graph Networks

### 2019-03-24 latex/iccv2019/main.tex
Final ICCV Submission (much last minute scramble)

%%%%%%%%% ABSTRACT
\begin{abstract}
    Image feature matching is a fundamental part of many geometric computer vision applications, and using multiple images can improve performance.
    In this work, we formulate multi-image matching as a graph embedding problem then use a Graph Neural Networks to learn an appropriate embedding function for aligning image features.
    We use cycle consistency to train our network in an unsupervised fashion, since ground truth correspondence is difficult or expensive to aquire.
    In addition, geometric consistency losses can be added at training time, even if the information is not available in the test set, unlike previous approaches that optimize cycle consistency directly.
    To the best of our knowledge, no other works have used learning for multi-image feature matching.
    Our experiments show that our method is competitive with other optimization based approaches.
\end{abstract}

@@

However, these multi-view consistency algorithms struggle in distributed and robust settings.
Deep learning has revolutionized how image features are computed \cite{yi2016lift}.
In this paper, we want to leverage the power of deep representations in order to compute feature descriptors that are robust across multiple views.
% Typically when applied to image or feature matching, deep neural networks (DNNs) are trained with photometric losses, directly using pixel color values.
% However direct, pixel-based losses have are not robust to the many confounding factors in the environment, such as lighting and view-point changes.
% Using features designed to deal with these confounding factors improves robustness to them, and learning may improve robustness even further.

@@

%------------------------------------------------------------------------
\begin{figure*}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures-CycleConsistencyMainFigure-v2.pdf}
\end{center}
  \caption{
    An illustration of the approach of this work.
    The Graph Neural Neural Network (GNN) \cite{battaglia2018relational} takes as input the graph of matches and then outputs a low rank embedding of the adjacency matrix of the graph.
    The GNN operates on an embedding over the vertices of the graph.
    In the figure, the GNN vertex embeddings are represented by different colors.
    The final embedding is used to construct a pairwise similarity matrix, which we train to be a low dimensional cycle-consistent representation of the graph adjacency matrix, thus pruning the erroneous matches.
    We train the network using a reconstruction loss on the similarity matrix with the noisy adjacency matrix, and thus do not need ground truth matches.
    In addition, we can use geometric consistency information, such as epipolar constraints, to assist training the network.
  }
\label{fig:pipeline}
\label{fig:onecol}
\end{figure*}

Unfortunately, there are obstacles to applying multi-view constraints directly to deep learning. 
To train networks, we need a large amount of labeled data.
In the case of multi-image feature matching, one would need hand labeled point correspondences between images, which are difficult and expensive to obtain.
Multi-view constraints are formulated in terms of sparse features, which traditional convolutional neural nets are not designed to handle.
Additionally, in the case of multi-image feature matching, geometric constraints would be helpful in rejecting outlier matches.

In this work, we propose to solve these problems using a Graph Neural Networks (GNNs) to operate on the correspondence graph.
The proposed method works directly on the correspondence graph, which is agnostic to how the correspondences were computed, thus allowing the algorithm to work in a broad class of environments.
To the best of our knowledge this work is the first to apply deep learning to the multi-view feature matching problem.
We use an unsupervised loss - the cycle consistency loss - to train the network, and thus avoiding the difficulty of expensive hand labeling.
% We incorporate geometric consistency losses to guide training, which helps training even when the geometric information is not available at test time.
Geometric consistency losses can aid training, even if such information is not available at test time.
Although our network is simple, it shows promising results compared to baselines which optimize for cycle-consistency without learned embeddings.
Furthermore, since inference requires only a single forward pass over the neural network, our approach is faster (to achieve comparable accuracy) than methods which must solve an optimization problem every time.
Our contributions are:
\begin{itemize}
\item We use a novel architecture to address the multi-image feature matching problem using GNNs with graph embeddings.
\item We introduce an unsupervised cycle consistency loss that does not require labeled correspondences to train.
\item We demonstrate the effectiveness of geometric consistency losses in improving training.
\item We perform experiments on the Rome16K \cite{li2010location} dataset to test the effectiveness of our method compared to optimization based methods.
\end{itemize}

@@

\subsection{Feature Matching}
Matching has a rich history of research in computer vision.
The most well known and widely used method for image matching is RANSAC \cite{fischler1981random}.
RANSAC, combined with hand-crafted feature descriptors such as SIFT \cite{lowe2004distinctive}, SURF \cite{bay2006surf}, BRIEF \cite{calonder2012brief}, or ORB \cite{mur2015orb}, has constituted the bulk of the matching literature for the last 40 years.
Finally, graph matching \cite{suh2015subgraph, hu2016distributable} can be used as a final step for more robust matches.

@@

\subsection{Deep Learning for Matching}
Previous attempts to improve image matching techniques using machine learning have focused on learning the descriptors given ground truth correspondence from curated datasets \cite{zagoruyko2015learning, yi2016lift, brachmann2017dsac}.
This approach is limited if one do not have the ability to get the ground truth correspondences.
There are other methods to build correspondences such as Choy et al.~\cite{choy2016universal}, but they only handle two-view constraints and require dense correspondences.
Most similar to our work, Yi et al.~\cite{yi2018learning} attempts to improve correspondences by learning match probabilities for RANSAC for greater robustness and speed.
However, they only focus on two view matching and do not exploit the advantages of the correspondence structure.
Note that while Zhu et al.~\cite{zhu2017unpaired} use cycle consistency in their loss, their method is restricted to pairwise cycle consistency, while we use multi-image cycle consistency.

@@

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.9\linewidth]{figures-UniverseOfFeatures-v2.pdf}
\end{center}
  \caption{
    Illustration of the universe of features.
    Each feature in each image corresponds to a 3D point in the scene.
    We can construct cycle consistent embeddings of the features by mapping each one to the one-hot vector of its corresponding 3D point.
    While there can be many features, there are fewer 3D points and thus this corresponds to a low rank factorization of the correspondence matrix.
    Best viewed in color.
  }
\label{fig:universefeatures}
\label{fig:onecol}
\end{figure}
Our goal is to learn optimal features that capture multiple image views by filtering out noisy feature matches.
The input to our algorithm is a set of features and noisy correspondences, and the output is a new set of features where the pairwise similarities of these features correspond to the true matches.
We do this by training the new set of feature embeddings to be cycle consistent.
We formulate this problem in terms of the correspondence graph of the features.
Vectors $\mat{x}$ and matrices $\mat{A}$ are denoted with boldface, and the $i^{th}$ row of a matrix is denoted $(\mat{A})_i$.
Graphs $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ have a set of vertices $\mathcal{V}$ and of edges $\mathcal{E} \subseteq \mathcal{V} \times \mathcal{V}$.
Implicitly in the subsequent discussions we assume the edges are directed.
For a vertex $v \in \mathcal{V}$ we use $\mathcal{N}^{h}(v)$ to denote the $h$-hop neighbors of $v$, with the superscript left out for 1-hop neighbors.
Similarly $\mathcal{E}(v)$ is used to denote the edges associated with $v$.
To denote the vertices connected to an edge $e \in \mathcal{E}$ we write $e(v_1, v_2)$.
\subsection{Correspondence Graph}
We assume there is an initial set of feature matches represented as a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$.
The graph is constructed from putative correspondences of image features across images, typically constructed using feature descriptor distance (e.g. SIFT feature distance).
While there are many interesting methods for computing these putative correspondences \cite{suh2015subgraph, yi2018learning}, we do not explore them in this work.
Typically putative correspondences are matched probabilistically, meaning a feature in one image matches to many features in another.
The ambiguity in the matches could come from repeated structures in the scene, insufficiently informative low-level feature descriptors, or just an error in the matching algorithm.
Filtering out these noisy matches is our primary learning goal.

Each vertex of the graph $v \in \mathcal{V}$ is an image feature, corresponding to some ground truth 3D point $\mat{P}(v)$.
Each edge $e = (v_1, v_2) \in \mathcal{E}$ is a potential correspondence.
Associated with each vertex $v$ is an embedding $\mat{f}_v \in \bR^{m_0}$, which can include the visual feature descriptor, position, scale, orientation, etc.
Similarly, each edge $e$ has an associated feature $\mat{f}_e \in \bR^{p_0}$ (in this work, initially just the weight of the feature association).
We use these features as the initialization for our learning algorithm.

@@

\subsection{Graph Neural Networks}
As input to our method we are given a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with the features described last section $\mat{f}_v \forall v \in \mathcal{V}$ and $\mat{f}_e \forall e in \mathcal{E}$.
As with any neural network, GNNs have layered outputs.
We describe the output of layer $k$ as $\mat{f}_v^{(k)} \forall v \in \mathcal{V}$ and $\mat{f}_e^{(k)} \forall e in \mathcal{E}$, with the initial embeddings denoted $\mat{f}_v^{(0)} = \mat{f}_v^{(0)}$ and $\mat{f}_e^{(0)} = \mat{f}_e$.

@@

Mathematically, this is expressed as: % TODO: Maybe make this MLP oriented?
\begin{align}
\mat{\tilde{f}}_{e(v_1,v_2)}^{(k+1)} =&\; \mat{U}_1^{(k)} \mat{f}_{e}^{(k)} + \mat{U}_1^{(k)} \mat{f}_{v_1}^{(k)} + \mat{U}_2^{(k)} \mat{f}_{v_2}^{(k)} \\
\mat{f}_{e(v_1,v_2)}^{(k+1)} =&\; \sigma \left(a^{(k)} + \mat{\tilde{f}}_{e(v_1,v_2)}^{(k+1)}\right) \\
\mat{\tilde{f}}_{v}^{(k+1)} =&\; \mat{W}_0^{(k)} \mat{f}_{v}^{(k)} + \sum_{e \in \mathcal{E}(v)} \mat{W}_1^{(k)} \mat{f}_{e}^{(k+1)} \\
\mat{f}_v^{(k+1)} =&\; \sigma\left(b^{(k)} + \mat{\tilde{f}}_v^{(k+1)}\right)
\end{align}
Here the learned weights are denoted $\mat{W}$ and $\mat{U}$, and the biases $a^{(k)}$ and  $b^{(k)}$.
In \cite{battaglia2018relational}, they allow for more sophisticated aggregation functions, but in this work we simply use the mean function.
In practice, we use MLPs between message passes between the vertices and edges for better expressiveness.

@@

Let $\mat{M}$ be the noiseless set of matches between our features, with $\mat{M}_{ij}$ being the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then we can say that, for all $i, j, k$:
\begin{equation}
\mat{M}_{ij} = \mat{M}_{ik} \mat{M}_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stay the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle consistency}, and has been used in a number of works to optimize for global consistency \cite{zhou2015multi, wang2017multi, leonardos2016distributed}.
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a matrix $\mat{X}_i$.
Then the cycle consistency becomes:
\begin{equation}
\mat{M}_{ij} = \mat{X}_{i}\mat{X}_{j}^\top
\label{eq:cycconsist2}
\end{equation}

@@

This reduces our complexity from $O(n^3)$ to $O(n^2)$.
We try to learn $\mat{E}$ to approximate $\mat{X}$ - in other words the final embedding should be an encoding of the universe of features. As we do not have the ground truth matches $\mat{M}$, we approximate it using the noisy adjacency matrix $\mat{A}$ of our correspondence graph. Thus our loss would be 
\begin{equation}
\mathcal{L}(\mat{A}, \mat{E}_n) = \mathcal{D}(\mat{A}, \mat{E}_n\mat{E}_n^\top)
\end{equation}
Here $\mathcal{D}$ could be an $L_2$ loss, $L_1$ loss, or many others. In this work, we use the $L_1$ loss. 
Note that because of this formulation, we can determine our embeddings only up to a rotation, as
\begin{equation}
\mat{E}R(\mat{E}R)^\top
= \mat{E}RR^\top\mat{E}^\top
= \mat{E}\mat{E}^\top
\label{eq:rotinvar}
\end{equation}
Thus when visualizing embeddings, we rotate them to make them more interpretable (see figure \ref{fig:embeddingsviz}).

@@

One of the main advantages of this approach over more traditional optimization based approaches is the ability to add geometric consistency information into the loss at training time, even if it is not available at test time.
The simplest way to add geometric consistency losses, and the approach we use here, is to use the epipolar constraint.
The epipolar constraint describes how the positions of features in different images corresponding to the same point should be related.
Given a relative pose $(R_{ij}, T_{ij})$ between two cameras $i$ and $j$  (transforms $j$ to $i$) the epipolar on corresponding feature locations $X_i$ and $X_j$:
\begin{equation}
X_{i}^\top \cross{T_{ij}}R_{ij} X_{j} = 0
\label{eq:essential_constraint_rel}
\end{equation}
In this work we use the two pose epipolar constraint \cite{tron2014quotient}:
\begin{equation}
X_{i}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{j} X_{j} = 0
\label{eq:essential_constraint}
\end{equation}
% Where the $(R_k, T_k)$ are the poses of cameras $i$ and $j$ respectively.
The constraint assumes that the $X_k$ are calibrated i.e. the camera intrinsics are known. 
Given the matrix of ground truth correspondences $\mat{M}_{ij}$ between camera $i$ and $j$, we can formulate this as a loss:
\begin{equation}
\mathcal{L}_{ij,geom}(\mat{M}_{ij}) = \sum_{k,l} (\mat{M}_{ij})_{kl} \left|X_{i,k}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{j} X_{j,l}\right|
\label{eq:geom_cost}
\end{equation}
For our purposes, since we use low rank embeddings $\mat{E}_{i}$, $\mat{E}_{j}$, the loss would read:
\begin{align}
\mathcal{L}_{geom}(\mat{E})
=&\; \mathrm{tr}(\mat{G}^\top \mat{E}\mat{E}^\top) = \sum_{i,j} (\mat{E})_{i} \cdot (\mat{E})_{j} (\mat{G})_{ij} \\
(\mat{G})_{ij} =&\; \left|X_{i}^\top R_{c(i)}^\top \cross{T_{c(j)} - T_{c(i)}}R_{c(j)} X_{j}\right| \nonumber
\label{eq:geom_cost2}
\end{align}
Where $c(i)$ is the appropriate camera for point index $i$.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
    Image feature matching is a fundamental part of many geometric computer vision applications, and using multiple images can improve performance.
    In this work, we formulate multi-image matching as a graph embedding problem then use a Graph Neural Network to learn an appropriate embedding function for aligning image features.
    We use cycle consistency to train our network in an unsupervised fashion, since ground truth correspondence is difficult or expensive to acquire.
    In addition, geometric consistency losses can be added at training time, even if the information is not available in the test set, unlike previous approaches which optimize cycle consistency directly.
    To the best of our knowledge, no other works have used learning for multi-image feature matching.
    Our experiments show that our method is competitive with other optimization based approaches.
\end{abstract}

@@

However, these multi-view consistency algorithms struggle in distributed and robust settings.
Having image features suited for this task would help improve performance, and deep learning has revolutionized how image features are computed \cite{yi2016lift}.
In this paper, we want to leverage the power of deep representations in order to compute feature descriptors that are robust across multiple views.
% Typically when applied to image or feature matching, deep neural networks (DNNs) are trained with photometric losses, directly using pixel color values.
% However direct, pixel-based losses have are not robust to the many confounding factors in the environment, such as lighting and view-point changes.
% Using features designed to deal with these confounding factors improves robustness to them, and learning may improve robustness even further.

@@

%------------------------------------------------------------------------
\begin{figure*}[t]
\begin{center}
  % \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
  \includegraphics[width=0.8\linewidth]{figures-CycleConsistencyMainFigure-v4.pdf}
\end{center}
  \caption{
    An illustration of the approach of this work.
    The Graph Neural Neural Network (GNN) \cite{battaglia2018relational} takes as input the graph of matches and then outputs a low rank embedding of the adjacency matrix of the graph.
    The GNN operates on an embedding over the vertices of the graph.
    In the figure, the GNN vertex embeddings are represented by different colors.
    The final embedding is used to construct a pairwise similarity matrix, which we train to be a low dimensional cycle-consistent representation of the graph adjacency matrix, thus pruning the erroneous matches.
    We train the network using a reconstruction loss on the similarity matrix with the noisy adjacency matrix, and thus do not need ground truth matches.
    In addition, we can use geometric consistency information, such as epipolar constraints, to assist training the network.
  }
\label{fig:pipeline}
\label{fig:onecol}
\end{figure*}

Unfortunately, there are obstacles to applying multi-view constraints directly to deep learning. 
To train networks, we need a large amount of labeled data.
In the case of multi-image feature matching, one would need hand labeled point correspondences between images, which are difficult and expensive to obtain.
Multi-view constraints are formulated in terms of sparse features, which traditional convolutional neural nets are not designed to handle.
Additionally, in the case of multi-image feature matching, geometric constraints would be helpful in rejecting outlier matches.
Geometric constraints, such as the epipolar constraint, can help disambiguate between visually similar features during training, and help the network learn better features robust to this kind of noise.

In this work, we propose to solve these problems using Graph Neural Networks (GNNs) to operate on the correspondence graph.
The proposed method works directly on the correspondence graph, which is agnostic to how the correspondences were computed, thus allowing the algorithm to work in a broad class of environments.
To the best of our knowledge this work is the first to apply deep learning to the multi-view feature matching problem.
We use an unsupervised loss - the cycle consistency loss - to train the network, and thus avoiding the difficulty of expensive hand labeling.
Geometric consistency losses can aid training, even if such information is not available at test time.
Although our network is simple, it shows promising results compared to baselines which optimize for cycle-consistency without learned embeddings, using a matrix factorization loss \cite{zhou2015multi, leonardos2016distributed}.
Furthermore, since inference requires only a single forward pass over the neural network, our approach is faster (to achieve comparable accuracy) than methods which must solve an optimization problem every time.
We perform experiments on the Rome16K \cite{li2010location} dataset to test the effectiveness of our method compared to optimization based methods.
Our contributions in this work are:
\begin{itemize}
\item We use a novel architecture to address the multi-image feature matching problem using GNNs with graph embeddings.
\item We introduce an unsupervised cycle consistency loss that does not require labeled correspondences to train.
\item We demonstrate the effectiveness of geometric consistency losses in improving training.
\end{itemize}

@@

\subsection{Feature Matching}
Matching has a rich history of research in computer vision.
Much work has be done using hand-crafted feature descriptors such as SIFT \cite{lowe2004distinctive}, SURF \cite{bay2006surf}, BRIEF \cite{calonder2012brief}, or ORB \cite{mur2015orb}.
RANSAC \cite{fischler1981random} is the most widely used robust estimation technique to filter out outliers from the matches.
The combination of RANSAC and hand-crafted feature descriptors has constituted the bulk of the matching literature for the last 40 years.
Finally, graph matching \cite{suh2015subgraph, hu2016distributable} can be used as a final step for more robust matches.

@@

\subsection{Deep Learning for Matching}
Previous attempts to improve image matching techniques using machine learning have focused on learning the descriptors given ground truth correspondence from curated datasets \cite{zagoruyko2015learning, yi2016lift, brachmann2017dsac}.
This approach is limited if one do not have the ability to get the ground truth correspondences.
There are other methods to build correspondences such as Choy et al.~\cite{choy2016universal}, but they only handle two-view constraints and require dense correspondences.
Most similar to our work, Yi et al.~\cite{yi2018learning} attempts to improve correspondences by learning match probabilities for RANSAC for greater robustness and speed.
However, they only focus on two view matching and do not exploit the advantages of the correspondence structure.
Note that while Zhu et al.~\cite{zhu2017unpaired} use cycle consistency in their loss, their method is restricted to pairwise cycle consistency (i.e. enforcing consistency by going back and forth between two images).
We use multi-image cycle consistency, which require consistency between 3 or more images.

@@

\begin{figure}[t]
\begin{center}
  \includegraphics[width=0.9\linewidth]{figures-UniverseOfFeatures-v2.pdf}
\end{center}
  \caption{
    Illustration of the universe of features.
    Each feature in each image corresponds to a 3D point in the scene.
    We can construct cycle consistent embeddings of the features by mapping each one to the one-hot vector of its corresponding 3D point.
    While there can be many features, there are fewer 3D points and thus this corresponds to a low rank factorization of the correspondence matrix.
    Best viewed in color.
  }
\label{fig:universefeatures}
\label{fig:onecol}
\end{figure}
Our goal is to learn optimal features that capture multiple image views by filtering out noisy feature matches.
An outline of our approach can be seen in figure \ref{fig:pipeline}.
The input to our algorithm is a set of features and noisy correspondences, and the output is a new set of features where the pairwise similarities of these features correspond to the true matches.
We do this by training the new set of feature embeddings to be cycle consistent.
We formulate this problem in terms of the correspondence graph of the features.
Vectors $\mat{x}$ and matrices $\mat{A}$ are denoted with boldface, and the $i^{th}$ row of a matrix is denoted $(\mat{A})_i$.
Graphs $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ have a set of vertices $\mathcal{V}$ and of edges $\mathcal{E} \subseteq \mathcal{V} \times \mathcal{V}$.
Implicitly in the subsequent discussions we assume the edges are directed.
For a vertex $v \in \mathcal{V}$ we use $\mathcal{N}^{h}(v)$ to denote the $h$-hop neighbors of $v$, with the superscript left out for 1-hop neighbors.
Similarly $\mathcal{E}(v)$ is used to denote the edges associated with $v$.
To denote the vertices connected to an edge $e \in \mathcal{E}$ we write $e(v_1, v_2)$.
\subsection{Correspondence Graph}
We assume there is an initial set of feature matches represented as a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$, with an associated adjacency matrix $\mat{A}$.
The graph is constructed from putative correspondences of image features across images, typically constructed using feature descriptor distance (e.g. SIFT feature distance).
While there are many interesting methods for computing these putative correspondences \cite{suh2015subgraph, yi2018learning}, we do not explore them in this work.
Typically putative correspondences are matched probabilistically, meaning a feature in one image matches to many features in another.
The ambiguity in the matches could come from repeated structures in the scene, insufficiently informative low-level feature descriptors, or just an error in the matching algorithm.
Filtering out these noisy matches is our primary learning goal.

Each vertex of the graph $v \in \mathcal{V}$ is an image feature, corresponding to some ground truth 3D point $\mat{P}(v)$.
Each edge $e = (v_1, v_2) \in \mathcal{E}$ is a potential correspondence.
Associated with each vertex $v$ is an embedding $\mat{f}_v \in \bR^m$, which can include the visual feature descriptor, position, scale, orientation, etc.
Similarly, each edge $e$ has an associated feature $\mat{f}_e \in \bR^p$ (in this work, initially just the weight of the feature association).
We use these features as the initialization for our learning algorithm.

@@

\subsection{Graph Neural Networks}
As input to our method we are given a graph $\mathcal{G} = (\mathcal{V}, \mathcal{E})$ with the features described last section $\mat{f}_v \forall v \in \mathcal{V}$ and $\mat{f}_e \forall e \in \mathcal{E}$.
As with any neural network, GNNs have layered outputs.
We describe the output of layer $k$ as $\mat{f}_v^{(k)} \in \bR^{m_k}\, \forall v \in \mathcal{V}$ and $\mat{f}_e^{(k)} \in \bR^{p_k}\, \forall e \in \mathcal{E}$, with the initial embeddings denoted $\mat{f}_v^{(0)} = \mat{f}_v$ and $\mat{f}_e^{(0)} = \mat{f}_e$.
It will be useful to write these features as matrices in later sections, so we denote the vertex embedding matrix as $\mat{F}_V^{(k)}$ and the edge embedding matrix as $\mat{F}_E^{(k)}$.
If a superscript is not specified then it refers to the final output of the network.

@@

Mathematically, this is expressed as: % TODO: Maybe make this MLP oriented?
\begin{align}
\mat{\tilde{f}}_{e(v_1,v_2)}^{(k+1)} =&\; \mat{U}_1^{(k)} \mat{f}_{e}^{(k)} + \mat{U}_1^{(k)} \mat{f}_{v_1}^{(k)} + \mat{U}_2^{(k)} \mat{f}_{v_2}^{(k)} \\
\mat{f}_{e(v_1,v_2)}^{(k+1)} =&\; \sigma \left(a^{(k)} + \mat{\tilde{f}}_{e(v_1,v_2)}^{(k+1)}\right) \\
\mat{\tilde{f}}_{v}^{(k+1)} =&\; \mat{W}_0^{(k)} \mat{f}_{v}^{(k)} + \sum_{e \in \mathcal{E}(v)} \mat{W}_1^{(k)} \mat{f}_{e}^{(k+1)} \\
\mat{f}_v^{(k+1)} =&\; \sigma\left(b^{(k)} + \mat{\tilde{f}}_v^{(k+1)}\right)
\end{align}
Here the learned weights are denoted $\mat{W}$ and $\mat{U}$, and the biases $a^{(k)}$ and  $b^{(k)}$.
In \cite{battaglia2018relational}, they allow for more sophisticated aggregation functions, but in this work we simply use the mean function.
In practice, we use MLPs in message passes between vertices and edges for better expressiveness

@@

Let $\mat{M}$ be the noiseless set of matches between our features, with $\mat{M}_{ij}$ being the matches between image $i$ and image $j$.
If the pairwise matches are globally consistent, then we can say that, for all $i, j, k$:
\begin{equation}
\mat{M}_{ij} = \mat{M}_{ik} \mat{M}_{kj}
\label{eq:cycconsist1}
\end{equation}
In other words, the matches between two images stay the same no matter what path is taken to get there. 
This constraint is known as \textit{cycle consistency}, and has been used in a number of works to optimize for global consistency \cite{zhou2015multi, wang2017multi, leonardos2016distributed}.
Stated in this form, there are $O(n^3)$ cycle consistency constraints to check.
A more elegant way to represent cycle consistency is to first create a `universe' of features that all images match to (see figure \ref{fig:universefeatures}).
Then, one can match the $i^{th}$ set of features to the universe using a ground-truth matching matrix $\mat{X}_i$.
Then the cycle consistency constraint becomes:
\begin{equation}
\mat{M}_{ij} = \mat{X}_{i}\mat{X}_{j}^\top
\label{eq:cycconsist2}
\end{equation}

@@

This reduces the number of our constraints from $O(n^3)$ to $O(n^2)$.
We try to learn vertex embeddings $\mat{F}_V$ to approximate $\mat{X}$ - in other words the final embedding should be an encoding of the universe of features.
As we do not have the ground truth matches $\mat{M}$, we approximate it using the noisy adjacency matrix $\mat{A}$ of our correspondence graph. Thus our loss would be 
\begin{equation}
\mathcal{L}(\mat{A}, \mat{F}_V) = \mathcal{D}(\mat{A}, \mat{F}_V \mat{F}_V^\top)
\end{equation}
Here $\mathcal{D}$ could be an $L_2$ loss, $L_1$ loss, or many others. In this work, we use the $L_1$ loss. 
Note that because of this formulation, we can determine our embeddings only up to a rotation, as
\begin{equation}
\mat{E}R(\mat{E}R)^\top
= \mat{E}RR^\top\mat{E}^\top
= \mat{E}\mat{E}^\top
\label{eq:rotinvar}
\end{equation}
Thus when visualizing embeddings, we rotate them to make them more interpretable (see figure \ref{fig:embeddingsviz}).

@@

One of the main advantages of this approach over more traditional optimization based approaches is the ability to add geometric consistency information into the loss at training time, even if it is not available at test time.
The simplest way to add geometric consistency losses, and the approach we use here, is to use the epipolar constraint.
The epipolar constraint describes how the positions of features in different images corresponding to the same point should be related.
An illustration of this is provided in figure \ref{fig:geoconsist}, showing how this loss can help reject erroneous points.
Given a relative pose $(R_{ij}, T_{ij})$ between two cameras $i$ and $j$  (transforms $j$ to $i$) the epipolar on corresponding feature locations $X_i$ and $X_j$:
\begin{equation}
X_{i}^\top \cross{T_{ij}}R_{ij} X_{j} = 0
\label{eq:essential_constraint_rel}
\end{equation}
In this work we use the two pose epipolar constraint \cite{tron2014quotient}:
\begin{equation}
X_{i}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{j} X_{j} = 0
\label{eq:essential_constraint}
\end{equation}
% Where the $(R_k, T_k)$ are the poses of cameras $i$ and $j$ respectively.
The constraint assumes that the $X_k$ are calibrated (i.e. the camera intrinsics are known). 
Given our vertex embeddings matrix $\mat{F}_V$, we can formulate a loss between all cameras $i$ and $j$:
\begin{align}
&\mathcal{L}_{ij,geom}(\mat{F}_V) = \nonumber \\ 
&\sum_{v \in \mathcal{V}(i),u \in \mathcal{V}(j)} \sum_{} (\mat{f}_v \cdot \mat{f}_u) \left|X_{v}^\top R_{i}^\top \cross{T_{j} - T_{i}}R_{c_j} X_{u}\right|
\label{eq:geom_cost}
\end{align}
With $\mathcal{V}(i)$ being the vertices associated with camera $i$.
For our purposes, since we use low rank embeddings $\mat{E}_{i}$, $\mat{E}_{j}$, the loss would read:
\begin{align} 
\mathcal{L}_{geom}(\mat{F}_V)
=&\; \mathrm{tr}(\mat{G}^\top \mat{F}_V\mat{F}_V^\top) = \sum_{k,l} (\mat{F}_V)_{k} \cdot (\mat{F}_V)_{l} (\mat{G})_{kl} \label{eq:geom_cost2} \\
(\mat{G})_{kl} =&\; \left|X_{k}^\top R_{c(k)}^\top \cross{T_{c(l)} - T_{c(k)}}R_{c(l)} X_{l}\right| \nonumber
\end{align}
Where $c(k)$ is the appropriate camera for point index $k$.

\begin{table*}
\begin{center}
\begin{tabular}{|l|c|c|c|c|}
\hline
Method (6 Views)                                     & $L_1$             & $L_2$             & Area under ROC    & Time (sec)        \\
\hline\hline
MatchALS \cite{zhou2015multi} 15 Iterations          & 0.101 $\pm$ 0.008 & 0.022 $\pm$ 0.004 & 0.918 $\pm$ 0.073 & 0.074 $\pm$ 0.008 \\ \hline
MatchALS \cite{zhou2015multi} 25 Iterations          & 0.059 $\pm$ 0.010 & 0.012 $\pm$ 0.004 & 0.921 $\pm$ 0.077 & 0.143 $\pm$ 0.064 \\ \hline
MatchALS \cite{zhou2015multi} 35 Iterations          & 0.046 $\pm$ 0.016 & 0.010 $\pm$ 0.005 & 0.910 $\pm$ 0.072 & 0.139 $\pm$ 0.041 \\ \hline
MatchALS \cite{zhou2015multi} 50 Iterations          & 0.029 $\pm$ 0.017 & 0.008 $\pm$ 0.005 & 0.905 $\pm$ 0.068 & 0.260 $\pm$ 0.048 \\ \hline
PGDDS0 \cite{leonardos2016distributed} 15 Iterations & 0.017 $\pm$ 0.002 & 0.007 $\pm$ 0.001 & 0.918 $\pm$ 0.087 & 0.796 $\pm$ 0.147 \\ \hline
PGDDS0 \cite{leonardos2016distributed} 25 Iterations & 0.016 $\pm$ 0.002 & 0.007 $\pm$ 0.002 & 0.919 $\pm$ 0.087 & 1.670 $\pm$ 0.328 \\ \hline
PGDDS0 \cite{leonardos2016distributed} 50 Iterations & 0.015 $\pm$ 0.002 & 0.006 $\pm$ 0.002 & 0.920 $\pm$ 0.087 & 3.363 $\pm$ 0.528 \\ \hline
Spectral \cite{pachauri2013solving}                  & 0.073 $\pm$ 0.006 & 0.027 $\pm$ 0.003 & 0.921 $\pm$ 0.083 & 0.036 $\pm$ 0.005 \\ \hline
\textbf{GNN (ours)}                                  & 0.044 $\pm$ 0.005 & 0.031 $\pm$ 0.005 & 0.872 $\pm$ 0.081 & 0.765 $\pm$ 0.046 \\ \hline

\hline\hline
Method (10 Views)                                   & $L_1$             & $L_2$             & Area under ROC    & Time (sec)        \\
\hline\hline
MatchALS \cite{zhou2015multi} 15 Iterations         & 0.114 $\pm$ 0.008 & 0.028 $\pm$ 0.004 & 0.915 $\pm$ 0.051 & 0.142 $\pm$ 0.009 \\ \hline
MatchALS \cite{zhou2015multi} 25 Iterations         & 0.092 $\pm$ 0.008 & 0.019 $\pm$ 0.003 & 0.912 $\pm$ 0.055 & 0.228 $\pm$ 0.014 \\ \hline
MatchALS \cite{zhou2015multi} 35 Iterations         & 0.065 $\pm$ 0.009 & 0.013 $\pm$ 0.003 & 0.907 $\pm$ 0.053 & 0.355 $\pm$ 0.073 \\ \hline
MatchALS \cite{zhou2015multi} 50 Iterations         & 0.045 $\pm$ 0.012 & 0.011 $\pm$ 0.004 & 0.914 $\pm$ 0.051 & 0.455 $\pm$ 0.022 \\ \hline
PGDDS \cite{leonardos2016distributed} 15 Iterations & 0.017 $\pm$ 0.001 & 0.008 $\pm$ 0.001 & 0.903 $\pm$ 0.061 & 1.225 $\pm$ 0.159 \\ \hline
PGDDS \cite{leonardos2016distributed} 25 Iterations & 0.016 $\pm$ 0.001 & 0.007 $\pm$ 0.001 & 0.904 $\pm$ 0.061 & 2.637 $\pm$ 0.357 \\ \hline
PGDDS \cite{leonardos2016distributed} 50 Iterations & 0.016 $\pm$ 0.001 & 0.007 $\pm$ 0.001 & 0.905 $\pm$ 0.061 & 6.116 $\pm$ 1.009 \\ \hline
Spectral \cite{pachauri2013solving}                 & 0.073 $\pm$ 0.005 & 0.029 $\pm$ 0.002 & 0.912 $\pm$ 0.057 & 0.081 $\pm$ 0.021 \\ \hline
\textbf{GNN (ours)}                                 & 0.053 $\pm$ 0.006 & 0.035 $\pm$ 0.005 & 0.872 $\pm$ 0.061 & 2.438 $\pm$ 0.070 \\ \hline

\end{tabular}
\end{center}

\caption{
Results on Rome16K Correspondence graphs, showing the mean and standard deviation of the $L_1$ and $L_2$.
Our method was not trained on ground truth correspondences but using unsupervised methods and geometric side losses.
As our method gives soft labels, we use cannot use precision or recall as is standard in testing cycle consistency \cite{zhou2015multi}.
Thus we test against ground truth correspondence graph adjacency matrices computed from the bundle adjustment output.
Our method performs better than 25 iteration of the MatchALS \cite{zhou2015multi} method, but does not perform as well as 50 iterations.
% Our network has only 12 layers, thus we have much greater efficiency per iteration.
We do not perform as well as the Projected Gradient Descent Doubly Stochastic (PGDDS) \cite{leonardos2016distributed} but we perform significantly faster than them.
Note that we perform much better in $L_1$ performance rather than $L_2$, as we optimized the network weights using an $L_1$ loss.
}
\label{tab:rome16ktab}
\end{table*}
